### 2020-06-03 code/TitanGPUmodel.Rmd
Reviewer 1 done

life_intervals =
  data.frame(days = as.numeric(diff(dates_life))/(60*60*24),
             year = factor(as.character(year(dates_life[-1])))) %>% filter(days >= 1) %>% 
  mutate(days = factor(as.integer(floor(days))))
ggp(ggplot(life_intervals, aes(days, fill = year)) + stat_count() +
      theme_bw(), file = "attention_intervals", height = 3)
```

>>>>

life_intervals =
  data.frame(days = as.numeric(diff(dates_life))/(60*60*24),
             year = factor(as.character(year(dates_life[-1])))) %>% filter(days >= 1) %>% 
  mutate(days = factor(as.integer(floor(days)), levels = 1:56))
ggp(ggplot(life_intervals, aes(days, fill = year)) + stat_count() +
      scale_x_discrete(breaks = seq(1, 56, 2), drop = FALSE) + theme_bw(),
    file = "attention_intervals", height = 3)
```

### 2020-06-03 dataprep.tex
Reviewer 1 done

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Data Collection and Preprocessing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Collection and Preprocessing}
\label{section:dataprep}
Data on Titan GPU life times is constructed from two sources:
inventory runs and failure event \rev{log} records. Two types of
failure events were collected \rev{for this study}: Double Bit Error
(DBE) and Off the Bus (OTB). \rev{While other types of errors were
  recorded in log files, as is reported in \cite{7266836}, these two
  kinds of errors became dominant in 2016 and were found to be the
  ``signature'' event of the GPU board failing resistor and a trigger
  for GPU replacement.}

An inventory of GPU serial numbers and their locations is recorded
each time the system boots. This typically occurs once every few days
but can be as frequent as a few times per day. A weekly inventory was
later added to run independently of system boots when they were fewer
and farther apart. A single inventory takes about a minute and is
recorded in a separate file. \rev{An inventory can be incomplete if a
  node response times out but such occurrences were relatively rare}.
Figure~\ref{fig:inventory} shows inter-inventory times that appear in
the data.
\begin{figure}[bt]
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/attention_intervals001.pdf}
  \end{center}
  \caption{Inter-inventory times, taken as time intervals between
    unique dates in the data and removing those less than a day.}
  \label{fig:inventory}
\end{figure}
Notably, there was a period of nearly two months and many periods of
over a week that Titan operated without a single reboot in the
2014-2015 time frame. \rev{However, by the time the weekly inventories
  were added, failure rates started to increase and lead to more
  frequent reboots and thus inventories.}

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Data Collection and Preprocessing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Collection and Preprocessing}
\label{section:dataprep}
Data on Titan GPU life times is constructed from two sources:
inventory runs and failure event \rev{log} records. Two types of
failure events were collected \rev{for this study}: Double Bit Error
(DBE) and Off the Bus (OTB). \rev{DBE is a memory error detected by
  parity bit checking, which can correct a single bit flip and detect
  but not correct a double bit flip. OTB is the loss of data swim lane
  connections to the GPU. While other types of errors were recorded in
  log files, as is reported in \cite{7266836}, OTB and DBE became
  dominant in 2016 and were found to be the ``signature'' event of the
  GPU board failing resistor and a trigger for GPU replacement.}

An inventory of GPU serial numbers and their locations was recorded
each time the system boots, \rev{typically for software updates and
  patches but sometimes for hardware swaps, although warm swaps of
  blades were usually possible.} Boots typically occured once every
few days but sometimes even a few times per day. A single inventory
takes about a minute and is recorded in a separate file. \rev{It can
  be incomplete if a node response times out but such occurrences were
  relatively rare.
\begin{figure}[bt]
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/attention_intervals001.pdf}
  \end{center}
  \caption{Inter-inventory times, taken as time intervals between
    unique dates in the data and removing those less than a day.}
  \label{fig:inventory}
\end{figure}
Figure~\ref{fig:inventory} shows that during 2014 and 2015,
inventories were far apart, even 56 days once in 2014. The purpose is
to illustrate that exact times for GPU removal from service are
``censored'' and we only get the inventory time information. We discuss
censoring and its implications in Sec.~\ref{section:survival}.}

### 2020-06-03 survival.tex
Reviewer 1 done

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Survival Analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \includegraphics[width=7in]{figs/km_cage-node_a001.pdf}
  \caption{Comparison of the old and new batches, including survival
    differences based on {\tt cage} and {\tt node} GPU locations.}
  \label{fig:km-all-cage-node}
\end{figure*}
\section{Survival Analysis}
\label{section:survival}
Survival analysis (SA) methods
\cite{survival-book}, 
sometimes referred to as time to event methods, focus on the effect of
the machine on the GPU units. These methods have roots in biomedical
statistics, where the effect on the patient is most important. So here
we treat the GPUs as patients to determine factors that contribute to
failures. SA methods use and combine information across the
operational lifetimes of all GPUs. For these analyses,
we take apart the location string of each GPU into variables {\em col,
  row, cage, slot, and node} and study the influence of the locations
on the GPU life times. The construction of a GPU life time is more
complex than it initially appears because the units are observed only
at reboot time, because most units were proactively replaced to
prevent failure, and because some units continue in operation after
OTB and DBE events (when a second reboot may be successful).

@@

Kaplan-Meier survival analysis (KM)
\cite{kaplan-meier1958,survival-book} starts with computing the
probability of survival beyond a given time. It is a nonparametric
technique that makes no specific failure model assumptions, such as
Weibull, Exponential, etc. The technique is able to use censored
observations and can also split the data into subpopulations to
compute separate survival curves.

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} \cite{survival-package} and
\pkg{survminer} \cite{survminer-package} for the KM analysis, which is
reported in \rev{Fig.}~\ref{fig:km-all-cage-node}. Within each {\em
  batch}, separate survival curves are computed for each {\em cage} by
{\em node} combination. Along with the survival curve estimate, this
analysis provides 95\% confidence region for survival probability
shaded around the curves.
% Inserting figure here so it appears at top of next page
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\em old}
    batch. \rev{All variables, {\em col}, {\em row}, {\em cage}, {\em
      slot}, and {\em node} are with respect to spatial locations
    shown in Fig.~\ref{fig:layout}}}
  \label{fig:cox-old}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_o_t001.pdf}
  \caption{\rev{GPU hazard ratios from Cox regression model on {\em
        old} batch, with the {\em col} variable in logical
      X-coordinate torus order. The torus order removes the peculiar
      interleaving pattern in Fig.~\ref{fig:cox-old} and presents a
      response that can be explained by a combination of system
      cooling and job scheduling.}}
  \label{fig:cox-torus}
\end{figure}

@@

To get more comparison power across the locations, we can use a
technique that in a sense averages over time. Cox proportional hazards
(CPH) regression analysis, can include covariates and estimates relative
risk averaged over time based on the covariates
\cite{Cox1972,Harrell2015}. The CPH regression function takes the form
\begin{displaymath}
  h(t) = h_0(t)e^{b_1 x_1 + b_2 x_2 \ldots b_k x_k},
\end{displaymath}
where $x_i$ are covariates, $h_0(t)$ is the {\em baseline hazard}, and
the $b_i$ are coefficients that measure the impact of the
covariates. The quantity $e^{b_i}$ is the hazard ratio (increased
average risk over baseline) for covariate $i$.

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Survival Analysis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \includegraphics[width=7in]{figs/km_cage-node_a001.pdf}
  \caption{Comparison of the old and new batches, including survival
    differences based on {\tt cage} and {\tt node} GPU locations.}
  \label{fig:km-all-cage-node}
\end{figure*}
\section{Survival Analysis}
\label{section:survival}
Survival analysis (SA) methods
\rev{\cite{survival-book}}, 
sometimes referred to as time to event methods, focus on the effect of
the machine on the GPU units. These methods have roots in biomedical
statistics, where the effect on the patient is most important. So here
we treat the GPUs as patients to determine factors that contribute to
failures. SA methods use and combine information across the
operational lifetimes of all GPUs. For these analyses,
we take apart the location string of each GPU into variables {\em col,
  row, cage, slot, and node} and study the influence of the locations
on the GPU life times. The construction of a GPU life time is more
complex than it initially appears because the units are observed only
at reboot time, because most units were proactively replaced to
prevent failure, and because some units continue in operation after
OTB and DBE events (when a second reboot may be successful).

@@

Kaplan-Meier survival analysis (KM)
\rev{\cite{kaplan-meier1958,survival-book}} starts with computing the
probability of survival beyond a given time. It is a nonparametric
technique that makes no specific failure model assumptions, such as
Weibull, Exponential, etc. The technique is able to use censored
observations and can also split the data into subpopulations to
compute separate survival curves.

If $T$ is the random variable of a GPU fail time, then its cumulative
distribution function $F(t) = Pr\{T < t\}$ gives the probability that
a GPU fails by duration $t$. The survival function is its complement
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t).
\end{displaymath}
It is the probability of being operational at duration $t$.  We use
the R packages \pkg{survival} \cite{survival-package} and
\pkg{survminer} \cite{survminer-package} for the KM analysis, which is
reported in \rev{Fig.}~\ref{fig:km-all-cage-node}. Within each {\em
  batch}, separate survival curves are computed for each {\em cage} by
{\em node} combination. Along with the survival curve estimate, this
analysis provides 95\% confidence region for survival probability
shaded around the curves.
% Inserting figure here so it appears at top of next page
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\em old}
    batch. \rev{All variables, {\em col}, {\em row}, {\em cage}, {\em
      slot}, and {\em node} are with respect to spatial locations
    shown in Fig.~\ref{fig:layout}. Notably, the {\em col} are
    physical columns of cabinets.}}
  \label{fig:cox-old}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/cox_o_t001.pdf}
  \caption{\rev{GPU hazard ratios from Cox regression model on {\em
        old} batch, with the {\em col} variable in logical
      X-coordinate torus order. The torus order removes the peculiar
      interleaving pattern in Fig.~\ref{fig:cox-old} and presents a
      response that can be explained by an interaction of system
      cooling and job scheduling.}}
  \label{fig:cox-torus}
\end{figure}

@@

To get more comparison power across the locations, we can use a
technique that in a sense averages over time. Cox proportional hazards
(CPH) regression analysis, can include covariates and estimates relative
risk averaged over time based on the covariates
\rev{\cite{Cox1972,Harrell2015}}. The CPH regression function takes the form
\begin{displaymath}
  h(t) = h_0(t)e^{b_1 x_1 + b_2 x_2 \ldots b_k x_k},
\end{displaymath}
where $x_i$ are covariates, $h_0(t)$ is the {\em baseline hazard}, and
the $b_i$ are coefficients that measure the impact of the
covariates. The quantity $e^{b_i}$ is the hazard ratio (increased
average risk over baseline) for covariate $i$.
